---
title: "Example with diabetes and stroke"
output: rmarkdown::html_vignette
description: >
    An example use case of the simulation framework for modeling the incidence of diabetes and stroke, and mortality.
vignette: >
    %\VignetteIndexEntry{Example with diabetes and stroke}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

In this example we consider a simulator for the incidence of type-2 diabetes and stroke with various risk factors as well as mortality from stroke and other causes.

## Data files
To run the example, the following data files are needed

 - The Finnish WHO MONICA data: http://www.thl.fi/publications/monica/monograph_cd/data/form04_3.zip
 - Finnish mortality statistics from Statistics Finland for the year 2017. The material was downloaded from Statistics Finland's interface service on 10 May 2020 with the license [CC BY 4.0](https://creativecommons.org/licenses/by/4.0/deed.en): http://users.jyu.fi/~santikka/Sima/data/kuol_007_201700.csv
 - Finnish age structure statistics from Statistics Finland for the year 2017. The material was downloaded from Statistics Finland's interface service on 11 May 2020 with the license [CC BY 4.0](https://creativecommons.org/licenses/by/4.0/deed.en): http://users.jyu.fi/~santikka/Sima/data/vaesto_3112_2017.csv
 
The data files (with MONICA data unzipped) should be placed to directory called `data` in relation to the working directory for the purposes of this example.

## R scripts
The R script files required to run the example can be downloaded at http://users.jyu.fi/~santikka/Sima/DiabetesAndStroke.zip  

The zip file contains the following

 - `DiabetesAndStroke.R` is the primary script that contains the all the simulator definition and simulation runs.
 - `DiabetesAndStrokeEvents.R` contains all the event definitions.
 - `DiabetesAndStrokeData.R` contains the function to generate the initial population.
 - `DiabetesAndStrokeFunctions.R` contains utility function definitions that are used throughout.
 
The contents of these scripts are also listed in the following sections for direct viewing
 
## Main script

The file `DiabetesAndStroke.R` provides the code to run the entire example

```{r, eval = FALSE}
# An example use case of the simulator for 
# modelling the occurrence of diabetes and stroke
# in a synthetic population with Finnish population
# characteristics in the year 2017.

# Uncomment to install packages
# install.packages(c("data.table", "ggplot2", "bestNormalize", "devtools"))
# devtools::install_github("santikk/Sima")

library(Sima)
library(data.table)
library(ggplot2)


# The file_path should be set to the location of the R-scripts, such that
# the data-directory is present in this location

file_path <- ""
setwd(file_path)

# Verify that all data files are available
if (!file.exists("data/vaesto_3112_2017.csv")) {
    stop("Data file vaesto_3112_2017.csv not found in data directory.")
}

if (!file.exists("data/kuol_007_201700.csv")) {
    stop("Data file kuol_007_201700 not found in data directory.")
}

if (!file.exists("data/form04_3.txt")) {
    stop("Data file form04_3.txt not found in data directory.")
}

#####################
# Utility Functions #
#####################

source("DiabetesAndStrokeFunctions.R")

##########
# Events #
##########

source("DiabetesAndStrokeEvents.R")

######################
# Initial Population #
######################

source("DiabetesAndStrokeData.R")

##################
# Initialization #
##################

sim <- Simulator$new(
    initializer = initializer,
    manipulation_events = all_events,
    time = Time$new(unit = "day", start = Sys.Date()),
    seeds = 242,
    n = 1e6
)

###############
# Calibration #
###############

## Uncomment to enable a parallel cluster for remaining computations
# if (require(doParallel)) {
#    nc <- parallel::detectCores()
#    cl <- parallel::makeCluster(nc)
#    doParallel::registerDoParallel(cl)
#    ## User made functions/objects have to be manually exported, list them in 'export'
#    sim$start_cluster(cl, nc, export = c("logit", "expit", "rbern"))
# }

# Uncomment to perform calibration
#calib_result <- optim(c(18, 140), mortality_objective, method = "Nelder-Mead")

# Using a population of 1 million individuals, the following optimum is found:
# c(13.51056, 161.12780, 5.41859)

# Configure the simulator to use the optimal parameters (if calibration not run)
# Here, only 1st parameter is configured for stroke mortality
sim$reconfigure(list("Other Mortality" = c(13.51056, 161.12780), "Stroke Mortality" = list(5.41859, 1)))

# Visualize and compare simulated mortality to official statistics
sim$run(364)
sim$stop_cluster()

mortality <- read.csv(file = "data/kuol_007_201700.csv", header = TRUE, sep = ",")
mortality[,3] <- mortality[,3]/1000
n <- nrow(mortality)
mortality <- mortality[mortality$Age >= 30,]
calib_result <- mortality_df(sim$get_population(), min_age = 30)
sim_mortality <- calib_result$out

# Proportion of fatal stroke deaths of all deaths
sim_stroke <- calib_result$stroke_prop

df <- data.frame(age = rep(sim_mortality$age_group, 2),
                 mortality = c(sim_mortality$deaths/sim_mortality$size, mortality[,3]),
                 group = gl(2, nrow(mortality), labels = c("Simulated", "Reported")))

# Plot comparing mortality in the synthetic population to the official statistics
g <- ggplot(df, aes(x = age, y = mortality, linetype = group)) + 
    geom_line(lwd = 0.5) +
    theme_bw(base_size = 16) +
    theme(
        legend.position = c(0.190, 0.85),
        legend.key.width = unit(1.75, "cm"),
        legend.key.size = unit(1.5, "line"),
        legend.text = element_text(size = 14),
        plot.margin = unit(c(5.5, 12, 5.5, 5.5), "points"),
        legend.background = element_rect(fill = "white", size = 2, linetype = "solid")
    ) +
    scale_x_continuous(limits = c(30, 100), breaks = 3:10 * 10, expand = c(0,0)) +
    scale_y_continuous(limits = c(0, 0.4), expand = c(0.025, 0)) +
    scale_linetype_manual(name = NULL, values = c("longdash", "solid")) +
    labs(x = "Age", y = "Mortality", title = NULL)
g

######################
# Selective sampling #
######################

# Create a smaller sub population corresponding to the target sample 
# (increase population size to 3.6 million to include the 
#  entire population in the following comparisions [pop])
sim_sample <- Simulator$new(
    initializer = initializer,
    manipulation_events = all_events,
    time = Time$new(unit = "day", start = Sys.Date()),
    seeds = 242,
    n = 1e4
)

# Configure the simulator to use the optimal parameters
# Here, only 1st parameter is configured for stroke mortality
sim_sample$reconfigure(list("Other Mortality" = c(13.51056, 161.12780), "Stroke Mortality" = list(5.41859, 1)))

missingness <- function(dt) {
    beta <- c(-9.484330, -0.311690, 0.113614, -0.090696, 0.036383, 0.689770)
    X <- as.matrix(dt[ ,.(const = 1, sex, age, bmi, waist_circumference, smoking)])
    return(rbern(expit(as.numeric(X %*% beta))))
}

pop <- data.table::copy(sim_sample$get_population())
m <- min(1e4, nrow(pop))
baseline_sample <- pop$id %in% sample(1:m, m, replace = FALSE)
baseline_nonresp <- missingness(pop[baseline_sample,])
baseline_stroke <- pop[baseline_sample,stroke]

# Uncomment to enable a parallel cluster for remaining computations
#if (require(doParallel)) {
#    nc <- parallel::detectCores()
#    cl <- parallel::makeCluster(nc)
#    doParallel::registerDoParallel(cl)
#    ## User made functions/objects have to be manually exported, list them in 'export'
#    sim_sample$start_cluster(cl, nc, export = c("expit", "logit", "rbern"))
#}

sim_sample$run(10 * 365)
sim_sample$stop_cluster()

# Update stroke status after 10 years
pop_10y <- sim_sample$get_population()[order(id),]
pop[baseline_sample, stroke := pop_10y[baseline_sample,stroke]]

# Comparing two subsets: one with dropout (miss), one without (full)
# (and one with the entire population, if included, otherwise same as full)
included_miss <- baseline_sample & baseline_nonresp == 0 & baseline_stroke == 0
included_full <- baseline_sample & baseline_stroke == 0
#included_pop <- baseline_stroke

# Separate models for males and females
model_miss_m <- glm(stroke ~ age + smoking + blood_pressure + hdl_cholesterol + diabetes + parent_stroke, 
                    data = pop[included_miss & sex == 0, ], family = binomial())
model_miss_f <- glm(stroke ~ age + smoking + blood_pressure + hdl_cholesterol + diabetes + parent_stroke, 
                    data = pop[included_miss & sex == 1, ], family = binomial())

model_full_m <- glm(stroke ~ age + smoking + blood_pressure + hdl_cholesterol + diabetes + parent_stroke, 
                    data = pop[included_full & sex == 0, ], family = binomial())
model_full_f <- glm(stroke ~ age + smoking + blood_pressure + hdl_cholesterol + diabetes + parent_stroke, 
                    data = pop[included_full & sex == 1, ], family = binomial())

#model_pop_m <- glm(stroke ~ age + smoking + blood_pressure + hdl_cholesterol + diabetes + parent_stroke, 
#                    data = pop[included_pop & sex == 0, ], family = binomial())
#model_pop_f <- glm(stroke ~ age + smoking + blood_pressure + hdl_cholesterol + diabetes + parent_stroke, 
#                    data = pop[included_pop & sex == 1, ], family = binomial())

summary(model_miss_m)
summary(model_miss_f)
summary(model_full_m)
summary(model_full_f)
#summary(model_pop_f)
#summary(model_pop_m)

round(exp(coef(model_miss_m)), 2)
round(exp(coef(model_miss_f)), 2)
round(exp(coef(model_full_m)), 2)
round(exp(coef(model_full_f)), 2)
#round(exp(coef(model_pop_m)), 2)
#round(exp(coef(model_pop_f)), 2)

round(exp(confint(model_miss_m)), 2)
round(exp(confint(model_miss_f)), 2)
round(exp(confint(model_full_m)), 2)
round(exp(confint(model_full_f)), 2)
#round(exp(confint(model_pop_m)), 2)
#round(exp(confint(model_pop_f)), 2)
```

## Events

The file `DiabetesAndStrokeEvents.R` includes the event definitions

```{r, eval = FALSE}
#####################
# Event definitions #
#####################

aging <- ManipulationEvent$new(
    name = "Aging",
    description = "Normal aging of individuals",
    parameters = c(),
    mechanism = expression({
        status[living, age := age + 1.0 / 365.0]
    })
)

other_mortality <- ManipulationEvent$new(
    name = "Other Mortality",
    description = "Risk of death when no stroke has occurred",
    parameters = c(13.16346, 165.34514),
    mechanism = expression({
        status[living & stroke == 0L, alive := rbern(1 - pweibull(age, shape = par[1], scale = par[2]))]
    })
)

get_stroke <- ManipulationEvent$new(
    name = "Stroke",
    description = "Risk of suffering a stroke",
    parameters = c(
        # const  age    smoking  bp        hdl     diab.   parents' stroke
        -11.699, 0.1153, 0.4981, 0.0149,  -0.4406, 0.879,  0.2933, # stroke, male
        -7.966,  0.0633, 0.4163, 0.00893, -0.7636, 1.2383, 0.5470  # stroke, female
    ),
    mechanism = expression({
        subset_male   <- status[living & stroke == 0L & sex == 0L, which = TRUE]
        subset_female <- status[living & stroke == 0L & sex == 1L, which = TRUE]
        X_male   <- as.matrix(status[subset_male,   .(const = 1, age, smoking, blood_pressure, hdl_cholesterol, diabetes, parent_stroke)])
        X_female <- as.matrix(status[subset_female, .(const = 1, age, smoking, blood_pressure, hdl_cholesterol, diabetes, parent_stroke)])

        # Compute 1 day risk from 10 year risk by assuming a Poisson process.
        prob_male_10year   <- 1 - pmin(expit( (X_male %*% par[1:7])[ ,1] ), 1)
        prob_female_10year <- 1 - pmin(expit( (X_female %*% par[8:14])[ ,1] ), 1)
        lambda_male   <- -log(pmax(prob_male_10year, 1e-12)) / 3650
        lambda_female <- -log(pmax(prob_female_10year, 1e-12)) / 3650
        new_stroke_male   <- rbern(1 - exp(-lambda_male))
        new_stroke_female <- rbern(1 - exp(-lambda_female))

        # Assign dates for stroke occurrence, update stroke status
        status[subset_male,   `:=` (age_at_stroke = age * new_stroke_male,   stroke = new_stroke_male)]
        status[subset_female, `:=` (age_at_stroke = age * new_stroke_female, stroke = new_stroke_female)]
    })
)

get_diabetes <- ManipulationEvent$new(
    name = "Diabetes",
    description = "Risk of getting diabetes",
    parameters = c(
        # const  age(45-54) age(55-64) bmi(25-30) bmi(>30) w. circ. (1) w.circ. (2) bp med. h. glucose
        -5.514,  0.628,     0.892,     0.165,     1.096,   0.857,       1.350,      0.711,  2.139
    ),
    mechanism = expression({
        subset <- status[living & diabetes == 0L & age >= 45, which = TRUE]
        X <- as.matrix(status[subset, .(
            const = 1,
            age1 = (age >= 45 & age < 54), 
            age2 = (age >= 54), 
            bmi1 = (bmi > 25 & bmi <= 30), 
            bmi2 = (bmi > 30), 
            waist_circumference1 = (sex == 0L) * (waist_circumference >= 94 & waist_circumference < 102) +
                                   (sex == 1L) * (waist_circumference >= 80 & waist_circumference < 88),
            waist_circumference2 = (sex == 0L) * (waist_circumference >= 102) + 
                                   (sex == 1L) * (waist_circumference >= 88),
            takes_bp_medicine, 
            high_glucose)
        ])

        # Compute 1 day risk from 10 year risk by assuming a Poisson process.
        prob_10year <- 1 - pmin(expit((X %*% par)[ ,1]), 1)
        lambda <- -log(pmax(prob_10year, 1e-12)) / 3650
        new_diabetes <- rbern(1 - exp(-lambda))

        # Update diabetes status
        status[subset, `:=` (diabetes = new_diabetes)]
    })
)

stroke_mortality <- ManipulationEvent$new(
    name = "Stroke Mortality",
    description = "Risk of death after suffering a stroke",
    parameters = c(5.41859, 8.17473693, -2.47553590, -0.01651868),
    mechanism = expression({
        status[living & stroke == 1L, alive := rbern(ifelse(age - age_at_stroke < (28.0 / 365.0), par[1], expit(par[2] + par[3] * exp(((age - age_at_stroke) * 365 - 27) * par[4]))))]
    })
)

bp_medication <- ManipulationEvent$new(
    name = "Medicine",
    description = "Blood pressure medication",
    parameters = c(140),
    mechanism = expression({
        status[living & blood_pressure >= par[1] & takes_bp_medicine == 0L, `:=` (takes_bp_medicine = 1L, blood_pressure = blood_pressure - 5)]
    })
)

# Further event organizing
stroke_events  <- list(stroke_mortality, get_stroke)
attr(stroke_events, "ordered") <- TRUE
other_events <- list(stroke_events, other_mortality, bp_medication, get_diabetes)
all_events <- list(aging, other_events)
attr(all_events, "ordered") <- TRUE
```

## Initial population

The file `DiabetesAndStrokeData.R` contains the following function for generating the initial population (of size `n`)

```{r, eval = FALSE}
initializer <- function(n) {
    # Seed for initial data generation
    set.seed(100)
    dqrng::dqset.seed(100)
    widths <- c(2,1,2,2,6,1,3,8,8,1,1,1,1,2,1,3,1,1,4,1,3,2,1,3,1,3,2,1,1,1,1,1,1,1,1,1,2,1,
                1,3,3,2,3,3,2,1,2,2,4,2,3,3,8,3,3,8,3,4,2,3,4,4,4,2,1,2,1,1,1,1,1,4,4,4,4)
    colnames <- c("form","versn","centre","runit","serial","numsur","samunit","dexam","mbirth",
                  "agegr","sex","marit","edlevel","school","cigs","numcigs","daycigs","evercig",
                  "stop","iflyear","maxcigs","cigage","cigarsm","cigar","pipesm","pipe","othersm",
                  "hibp","drugs","bprecd","hich","chdt","chrx","chrecd","asp","menop","agem",
                  "horm","pill","syst1","diast1","rz1","syst2","diast2","rz2","cuff","arm",
                  "bpcoder","timebp","rtemp","chol","choldl","dchol","hdl","hdldl","dhdl","scn",
                  "cotin","carbmon","height","weight","waist","hip","whcoder","oversion","eage",
                  "eageg","cohort1","cohort2","edtert1","edtert2","systm","diastm","chola","hdla")
    monica <- read.fwf("data/form04_3.txt", widths = widths, col.names = colnames)

    age_structure <- read.csv2(file = "data/vaesto_3112_2017.csv", header = TRUE)
    sex <- rbin(n)
    mf <- age_structure[,2:3] 
    age_structure[,2:3] <- mf / colSums(mf)[col(mf)]
    age <- (sex == 0L) * sample(30:99, n, replace = TRUE, prob = age_structure[,2]) +
           (sex == 1L) * sample(30:99, n, replace = TRUE, prob = age_structure[,3])
    smoking <- sex_age_init(n, sex, age, 
                            c(0.319, 0.298, 0.278, 0.219, 0.109, 0.016),
                            c(0.229, 0.193, 0.184, 0.147, 0.054, 0.016), 30, 80, 10)

    parent_stroke <- rbin(n, prob = 0.127)
    high_glucose <- sex_age_init(n, sex, age, 
                                 c(0.013, 0.043, 0.080, 0.114, 0.128, 0.110), 
                                 c(0.019, 0.012, 0.030, 0.055, 0.121, 0.093), 30, 80, 10)
    stroke <- sex_age_init(n, sex, age, 
                           c(0.027, 0.063, 0.091, 0.163), 
                           c(0.013, 0.029, 0.061, 0.137), 50, 80, 10)
    age_at_stroke <- rep(0.0, n)
    age_at_stroke[stroke == 1L] <- pmax(age[stroke == 1L] - 
        (rweibull(sum(stroke), shape = 0.509246, scale = 1634.480968)) / 365.0, 30)

    # decoding missing values
    monica$height[monica$height == 999] <- NA
    monica$weight[monica$weight == 9999] <- NA
    monica$waist[monica$waist == 9999] <- NA
    monica$systm[monica$systm == 9999] <- NA
    monica$chol[monica$chol >= 888] <- NA
    monica$hdl[monica$hdl >= 777] <- NA

    # unit transformations
    monica$weight <- monica$weight / 10 # transform to kg
    monica$waist <- monica$waist / 10 # transform to cm
    monica$systm <- monica$systm / 10 # transform to mmHg
    monica$chol <- monica$chol / 10 # transform to mmol/l
    monica$hdl <- monica$hdl / 100 # transform to mmol/l
    monica$bmi <- monica$weight / (monica$height / 100)^2
    monica$sex <- monica$sex - 1 # female = 1
    #monica$smoking <- 0 + (monica$cigs %in% c(1,3))
    monica$smoking <- 0 + (monica$cigs %in% c(1, 3) | 
                           monica$cigarsm %in% c(1, 3) | 
                           monica$pipesm %in% c(1, 3))

    selcol <- c("bmi", "waist", "chol", "hdl", "systm")
    m <- length(selcol)
    monica2 <- monica[,c("eage", "sex", "smoking", selcol)]
    out <- c(which.min(monica2$waist), which.max(monica2$chol))
    monica2 <- monica2[-out,]
    monica2$agegr <- cut(monica2$eage, breaks = c(0, 30, 40, 50, 60, 70), right = FALSE)

    # Generate initial values in each subgroup of sex, smoking and age
    norm_grp <- expand.grid(sex = 0:1, smoking = 0:1, agegr = 1:5)
    k <- nrow(norm_grp)
    norm_dat <- vector(mode = "list", length = k)
    norm_par <- vector(mode = "list", length = k)

    for (i in 1:k) {
        ind <- (monica2$sex == norm_grp[i,1] & 
                monica2$smoking == norm_grp[i,2] & 
                as.numeric(monica2$agegr) == norm_grp[i,3])
        ind[is.na(ind)] <- FALSE
        norm_par[[i]]$trans <- list()
        norm_dat[[i]] <- matrix(0, sum(ind), m)
        for (j in 1:m) {
            bc <- bestNormalize::boxcox(monica2[ind,selcol[j]])
            norm_par[[i]]$trans[[j]] <- bc
            norm_dat[[i]][,j] <- bc$x.t
        }
        norm_par[[i]]$mu <- colMeans(norm_dat[[i]], na.rm = TRUE)
        norm_par[[i]]$Sigma <- cov(norm_dat[[i]], use = "complete.obs")
    }

    X <- matrix(0, n, m)

    age_mon <- as.numeric(cut(age, breaks = c(0, 30, 40, 50, 60, 100), right = FALSE))
    for (i in 5:k) {
        ind <- (sex == norm_grp[i,1] &
                smoking == norm_grp[i,2] & 
                age_mon == norm_grp[i,3])
        temp <- MASS::mvrnorm(sum(ind), mu = norm_par[[i]]$mu, Sigma = norm_par[[i]]$Sigma)
        for (j in 1:m) {
            temp[,j] <- predict(norm_par[[i]]$trans[[j]], temp[,j], inverse = TRUE)
        }
        X[ind,] <- temp
    }

    # Limits data to values that have been actually observed
    X <- impose_limits(X, sex, smoking, age_mon, 
                       list(c(16, 72.5), c(50, 152), c(2.5, 11.5), c(0.37, 3.62), c(78, 252)))

    bmi <- X[,1]
    waist_circumference <- X[,2]
    cholesterol <- X[,3]
    hdl_cholesterol <- X[,4]
    blood_pressure <- X[,5]

    # Generating diabetes status based on FINDRISC risk score
    diabetes_par <- c(-5.514, 0.628, 0.892, 0.165, 1.096, 0.857, 1.350, 2.139)
    D <- matrix(0, n, 8)
    D[,1] <- 1
    D[,2] <- (age >= 45 & age < 54)
    D[,3] <- (age >= 54)
    D[,4] <- (bmi > 25 & bmi <= 30)
    D[,5] <- (bmi > 30)
    D[,6] <- (sex == 0L) * (waist_circumference >= 94 & waist_circumference < 102) +
             (sex == 1L) * (waist_circumference >= 80 & waist_circumference < 88)
    D[,7] <- (sex == 0L) * (waist_circumference >= 102) + 
             (sex == 1L) * (waist_circumference >= 88)
    D[,8] <- high_glucose
    diabetes_risk <- expit(as.numeric(D %*% diabetes_par))
    diabetes_risk[sex == 0L] <- diabetes_risk[sex == 0L] / mean(diabetes_risk[sex == 0L]) * 0.15
    diabetes_risk[sex == 1L] <- diabetes_risk[sex == 1L] / mean(diabetes_risk[sex == 1L]) * 0.10
    diabetes <- rbern(diabetes_risk)

    return(data.table(
        alive = rep(1L, n),
        age = as.double(age),
        age_at_start = as.double(age),
        age_at_stroke = age_at_stroke,
        sex = sex,
        smoking = smoking,
        diabetes = diabetes,
        parent_stroke = parent_stroke,
        high_glucose = high_glucose,
        stroke = stroke,
        bmi = bmi,
        waist_circumference = waist_circumference,
        cholesterol = cholesterol,
        hdl_cholesterol = hdl_cholesterol,
        blood_pressure = blood_pressure,
        takes_bp_medicine = rep(0L, n),
        id = 1:n
    ))

}
```

## Utility functions

The file `DiabetesAndStrokeFunctions.R` provides a set of useful functions used in defining events and the initial data

```{r, eval = FALSE}
## Bernoulli distribution with probability of success (1) 'prob' (vectorized over prob)

rbern <- function(prob) {
    return(ifelse(dqrng::dqrunif(length(prob)) < prob, 1L, 0L))
}

## Uniformly random date between 'start' and 'end' dates (inclusive)

rdate <- function(start, end, ...) {
    return(sample(seq(as.Date(start), as.Date(end), by = "day"), ...))
}

## Random binary, where the probability of 1 is 'prob'

rbin <- function(size, prob = 0.5) {
    return(sample(c(0L, 1L), size = size, replace = TRUE, prob = c(1 - prob, prob)))
}

## The logistic link function

logit <- function(x) log(x) - log(1 - x)

## Inverse logistic link

expit <- function(x) (1 + exp(-x))^(-1)

## Function to generate binary starting values based on age-based proportions
## n        : number of values to generate
## sex      : vector of sex for each individual
## age      : vector of ages (in years)
## val_m    : proportions for males in each age group
## val_f    : proportions for females in each age group
## age_min  : youngest age group, no individual should have an age below this
## age_max  : oldest distinct age group, ages above this are considered a single group
## age_step : size of each age group

sex_age_init <- function(n, sex, age, val_m, val_f, age_min, age_max, age_step) {
    groups <- (age_max - age_min) %/% age_step + 1
    age_group <- pmin((age - age_min) %/% age_step + 1, groups)
    age_group[age_group < 1] <- NA
    p <- (age_group > 0) * ((sex == 0) * val_m[age_group] + (sex == 1) * val_f[age_group])
    p[is.na(p)] <- 0
    return(rbinom(n, 1, p))
}

## Function to resample values within an allowed range
## X       : a matrix of values
## sex     : a vector of sex indicators
## smoking : a vector of smoking indicators
## age     : a vector of age groups
## limits  : a list of allowed ranges for column of 'X'

impose_limits <- function(X, sex, smoking, age, limits) {
    age <- as.numeric(age)
    groups <- expand.grid(unique(sex), unique(smoking), unique(age))
    for (i in 1:nrow(groups)) {
        g <- sex == groups[i,1] & smoking == groups[i,2] & age == groups[i,3]
        for (j in 1:ncol(X)) {
            out <- X[g,j] < limits[[j]][1] | X[g,j] > limits[[j]][2]
            bound <- !out
            g_out <- g
            g_out[g] <- out
            g_bound <- g
            g_bound[g] <- !out
            X[g_out,j] <- sample(X[g_bound,j], sum(out), replace = TRUE)
        }
    }
    return(X)
}

## Objective function to calibrate total mortality to mortality of the Finnish population
## x : current vector of parameter values

mortality_objective <- function(x) {
    mortality <- read.csv(file = "data/kuol_007_201700.csv", header = TRUE, sep = ",")
    mortality[,3] <- mortality[,3]/1000
    n <- nrow(mortality)
    mortality <- rbind(mortality, c(100, 2017, 2 * mortality[n,3] - mortality[n-1,3])) # Impute 100-year-olds from 99
    p <- list("Other Mortality" = x[1:2], "Stroke Mortality" = list(x[3], 1))
    calib_res <- sim$configure(t_sim = 364, p = p, output = mortality_df, min_age = 30)
    out <- calib_res$out
    stroke_prop <- calib_res$stroke_prop
    s_obj <- 1e7
    if (stroke_prop > 0) s_obj <- 2000 * (log(stroke_prop) - log(0.07534936))^2
    mortality <- mortality[mortality$Age %in% out$age_group,]
    w <- rep(1, nrow(out))
    w[out$age_group >= 80] <- 100 # Weight for mortalities >= 60, fit should be better here
    valid <- (out[ ,2] != 0 & out[ ,1] != out[,2] & out[ ,1] != 0)
    all_dead <- out[ ,1] == out[ ,2]
    all_alive <- out[ ,1] == 0
    obj <- sum(w[valid] * (log(mortality[valid,3]) - log(out[valid,1]/out[valid,2]))^2)
    obj <- obj + sum(w[all_dead] * log(mortality[all_dead,3])^2)
    obj <- obj + sum(w[all_alive] * log(mortality[all_alive,3])^2)
    return(obj + s_obj)
}

## A function used to calibrate total mortality
## dt        : the status data.table
## min_age   : the youngest age group

mortality_df <- function(dt, min_age = 0) {
    dt[ ,age_group := floor(age)]
    deaths <- dt[alive == 0, .N, by = age_group]
    size <- dt[ , .N, by = age_group]
    groups <- 0:max(dt$age_group)
    ages <- length(groups)
    out <- data.frame(deaths = rep(0, ages), size = rep(0, ages), age_group = groups)
    out[size$age_group + 1,"size"] <- size$N
    if (any(deaths$N > 0)) {
        out[deaths$age_group + 1,"deaths"] <- deaths$N
    }
    out <- out[out$age_group >= min_age, ]
    stroke_prop <- nrow(dt[alive == 0 & has_cardio == 1 & (age - cardio_disease) < (28/365),]) / nrow(dt[alive == 0,])
    return(list(out = out, stroke_prop = stroke_prop))
}
```

