---
title: "Example with diabetes and stroke"
output: rmarkdown::html_vignette
description: >
    An example use case of the simulation framework for modeling the incidence 
    of diabetes and stroke, and mortality with interventions
vignette: >
    %\VignetteIndexEntry{Example with diabetes and stroke}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

In this example we consider a simulator for the incidence of type-2 diabetes and stroke with various risk factors as well as mortality from stroke and other causes. Interventions that aim to reduce systolic blood pressure in the population are also considered

## Data files
To run the example, the following data files are needed

 - The Finnish WHO MONICA data: http://www.thl.fi/publications/monica/monograph_cd/data/form04_3.zip
 - Finnish mortality statistics from Statistics Finland for the year 2017. The material was downloaded from Statistics Finland's interface service on 10 May 2020 with the license [CC BY 4.0](https://creativecommons.org/licenses/by/4.0/deed.en): http://users.jyu.fi/~santikka/Sima/data/kuol_007_201700.csv
 - Finnish age structure statistics from Statistics Finland for the year 2017. The material was downloaded from Statistics Finland's interface service on 11 May 2020 with the license [CC BY 4.0](https://creativecommons.org/licenses/by/4.0/deed.en): http://users.jyu.fi/~santikka/Sima/data/vaesto_3112_2017.csv
 
The data files (with MONICA data unzipped) should be placed to directory called `data` in relation to the working directory for the purposes of this example.

## R scripts
The R script files required to run the example can be downloaded at http://users.jyu.fi/~santikka/Sima/diabetes_and_stroke.zip  

The zip file contains the following

 - `diabetes_and_stroke.R` is the primary script that contains the all the simulator definition and simulation runs.
 - `events.R` contains all the event definitions.
 - `data.R` contains the function to generate the initial population.
 - `functions.R` contains utility function definitions that are used throughout.
 
The contents of these scripts are also listed in the following sections for direct viewing
 
## Main script

The file `diabetes_and_stroke.R` provides the code to run the entire example

```{r, eval = FALSE, code = xfun::read_utf8('../../../examples/diabetes_and_stroke/diabetes_and_stroke.R')}
```

## Events

The file `events.R` includes the event definitions

```{r, eval = FALSE, code = xfun::read_utf8('../../../examples/diabetes_and_stroke/events.R')}
```

## Initial population

The file `data.R` contains the following function for generating the initial population (of size `n`)

```{r, eval = FALSE, code = xfun::read_utf8('../../../examples/diabetes_and_stroke/data.R')}
```

## Utility functions

The file `functions.R` provides a set of useful functions used in defining events and the initial data

```{r, eval = FALSE, code = xfun::read_utf8('../../../examples/diabetes_and_stroke/functions.R')}
```

