---
title: "Sick-sicker model"
output: rmarkdown::html_vignette
description: >
    In this example we re-implement the sick-sicker model that appears in
    Krijkamp EM, Alarid-Escudero F, Enns EA, Jalal HJ, Hunink MGM, Pechlivanoglou P. 
    Microsimulation modeling for health decision sciences using R: 
    A tutorial. Med Decis Making. 2018;38(3):400–22.
vignette: >
    %\VignetteIndexEntry{Sick-sicker model}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

In this example we consider a simple scenario where a person can occupy
one of four states: healthy, sick, sicker and dead. At each iteration,
we compute the transition probabilities in the population and change the
state accordingly. In addition, there are costs and quality adjusted life years 
associated with each state.

For more information and details about the model, please refer to the original article:
Krijkamp EM, Alarid-Escudero F, Enns EA, Jalal HJ, Hunink MGM, Pechlivanoglou P. 
Microsimulation modeling for health decision sciences using R: A tutorial.
*Med Decis Making.* 2018;**38(3)**:400–22.

The code used for comparative performance purposes can be found on GitHub:
https://github.com/DARTH-git/Microsimulation-tutorial

## R-code
Here we provide the R-code for implementing the model in the Sima framework.
We also see that the performance of Sima in this example is superior.

```{r, eval = FALSE, code = xfun::read_utf8('../../../examples/sick_sicker_comparison/sick_sicker_comparison.R')}
```
